\sys: Light-touch Data-Parallel Shell Processing

### 2020-10-09 paper/draft.tex
cdot in 4.3 and nits

% Another important feature is synchronization.
Different pipeline stages process data concurrently and possibly at different rates---\eg the second \ttt{curl} produces output at a significantly slower pace than the \ttt{grep} commands before and after it.
The \unix kernel facilitates scheduling, communication, and synchronization behind the scenes.

\begin{figure}[t]
\centering
\begin{minted}[fontsize=\footnotesize]{bash}
base="ftp://ftp.ncdc.noaa.gov/pub/data/noaa";
for y in {2015..2020}; do
 curl $base/$y | grep gz | tr -s" " | cut -d" " -f9 |
 sed "s;^;$base/$y/;" | xargs -n 1 curl -s | gunzip |
 cut -c 89-92 | grep -iv 999 | sort -rn | head -n 1 |
 sed "s/^/Maximum temperature for $y is: /"
done
\end{minted}
% \includegraphics[width=0.47\textwidth]{\detokenize{./figs/dish_example.pdf}}
\caption{
  \textbf{Calculating maximum temperatures per year.}
  The script downloads daily temperatures recorded across the U.S. for the years 2015--2020 and extracts the maximum for every year.
}
\vspace{-15pt}
\label{fig:example}
\end{figure}

@@

\begin{figure}[t]
\centering
% \includegraphics[width=0.49\textwidth]{\detokenize{./figs/dish_example2.pdf}}
\begin{minted}[fontsize=\footnotesize,escapeinside=()]{bash}
mkfifo $t{0,1...}
curl $base/$y > $t0 & cat $t0 | split $t1 $t2 &
cat $t1 | xargs -n1 curl -s >$t3 & 
cat $t2 | xargs -n1 curl -s >$t4 &
...
cat $t9 | sort -rn > $t11 & cat $t10 | sort -rn > $t12 &
cat $t11 | eager > $t13 & cat $t12 | eager > $t14 &
sort-merge -rn $t13 $t14 > $t15 &
cat $t15 | head -n1 > $out1 &
wait $! && get-pids | xargs -n 1 kill -SIGPIPE
\end{minted}
\caption{
  \textbf{Output of \ttiny{pash --width=2} for Fig.~\ref{fig:example} (fragment).}
  \sys orchestrates parallel execution through named pipes, parallel operators, and custom runtime primitives---\eg \ttiny{eager}, \ttiny{sort-merge}, and \ttiny{get-pids}.
}
\vspace{-15pt}
\label{fig:example2}
\end{figure}

@@

\heading{Stateless Commands}
The first class, \sta, contains commands that operate on individual line elements of their input, without maintaining state across invocations.
These are commands that can be expressed as a purely functional $map$ or $filter$---\eg \ttt{grep} filters out individual lines and \ttt{basename} removes a path prefix from a string.
They may produce multiple elements---\eg \ttt{tr} may insert {\sc NL} tokens---but always return empty output for empty input.
Workloads that use only stateless commands are trivial to parallelize:
  they do not require any synchronization to maintain correctness, nor caution about where to split inputs.
 % map :: (a -> b) -> [a] -> [b]

@@

\heading{Non-parallelizable Pure Commands}
The third class, \npu, contains commands that, while purely functional, cannot be parallelized.
This is because their internal state depends on prior state in the same pass in non-trivial ways. % should we say something about state machines?
For example, hashing commands such as \ttt{sha1sum} maintain complex state that has to be updated sequentially.
If parallelized on a single input, each stage would need to wait on the results of all previous stages foregoing any parallelism benefits.

@@

Similarly, some pure commands (such as \ttt{sort} and \ttt{wc}) can be
parallelized using divide-and-conquer techniques. More formally, these
pure commands $f$ can be implemented as a combination of a function map
$m$ and an associative aggregate $agg$, satisfying the following equation:
%
\[
\forall x, x', s, f(x \cdot x', s) = agg(m(x, s) \cdot m(x', s), s)
\]
%
\noindent
This means that we get the same output by applying $f$ to a
concatenation of two inputs $x, x'$ as by applying the aggregation function
$agg$ to the concatenation of the outputs produced by applying map $m$
to each of $x$ and $x'$.
%% \km{We had also discussed here the case of
%%   operators that can be described as associative and commutative
%%   functions on the data (e.g., sum). In this case the parallelization
%%   has the flavor of data parallelism: one can split the input stream
%%   in an arbitrary way (e.g., randomly or using some other
%%   load-balancing scheme) and merge the outputs. For the more broad
%%   class of associative operations (e.g., sort), the formulation of
%%   this paragraph makes perfect sense. I just want to mention that this
%%   feels a lot more like \textbf{batching} than map-reduce-like data
%%   parallelization.} \kk{Thanks, I think you are right. Batching is a
%%   better name for it. In the end we didn't handle
%%   commutative-associative function (that are more like map-reduce) and
%%   so we changes the name to map-fold to not imply commutativity.}

@@

$T$ can be extended straightforwardly to nodes $v \in$ \pur,
implemented by a map-aggregate pair $(m, agg)$ as $ v(x_1 \cdot x_2
\cdots x_n, s) \Rightarrow agg(m(x_1, s) \cdot m(x_2, s) \cdots m(x_n,
s), s)$.
As long as the pair $(m, agg)$ meets the three invariants outlined earlier~\sx{ext}, 
$T$ can be shown to be behavior-preserving.
% \todo{It is shown in Fig.}\ref{fig:parallelization-transformation}.

\heading{Auxiliary Transformations}
%
\sys also performs a set of auxiliary transformations $t_{1-3}$that are
depicted in \Cref{fig:auxiliary-transformations}. If a node has many
inputs, $t_1$ concatenates these inputs by inserting a \ttt{cat} node
to enable the parallelization transformations. In cases where a
parallelizable node has one input and is not preceded by a
concatenation, $t_2$ inserts a \ttt{cat} node that is preceded by its
inverse \ttt{split}, so that the concatenation can be commuted with
the node. Transformation $t_3$ inserts a relay node that
performs the identity transformation. Relay nodes can be useful for
monitoring and debugging, as well as for performance
improvements~\sx{optimizer}.
The additional relay nodes inserted by the auxiliary transformations affect the resulting % (task) 
  parallelism.

@@

% \subsection*{Acknowledgements}
% % Dumping people so that we don't forget
% We want to thank 
%   Andr\'e DeHon,
%   Ben Karel,
%   Jean-S\'ebastien L\'egar\'e, and 
%   Nick Roessler.
% This material is based upon work supported by DARPA contract no. HR00112020013 (ALRSCC) and NSF award CCF 1763514.

>>>>

% Another important feature is agg(synchronization.
Different pipeline stages process data concurrently and possibly at different rates---\eg the second \ttt{curl} produces output at a significantly slower pace than the \ttt{grep} commands before and after it.
The \unix kernel facilitates scheduling, communication, and synchronization behind the scenes.

\begin{figure}[t]
\centering
\begin{minted}[fontsize=\footnotesize]{bash}
base="ftp://ftp.ncdc.noaa.gov/pub/data/noaa";
for y in {2015..2019}; do
 curl $base/$y | grep gz | tr -s" " | cut -d" " -f9 |
 sed "s;^;$base/$y/;" | xargs -n 1 curl -s | gunzip |
 cut -c 89-92 | grep -iv 999 | sort -rn | head -n 1 |
 sed "s/^/Maximum temperature for $y is: /"
done
\end{minted}
% \includegraphics[width=0.47\textwidth]{\detokenize{./figs/dish_example.pdf}}
\caption{
  \textbf{Calculating maximum temperatures per year.}
  The script downloads daily temperatures recorded across the U.S. for the years 2015--2019 and extracts the maximum for every year.
}
\vspace{-15pt}
\label{fig:example}
\end{figure}

@@

\begin{figure}[t]
\centering
% \includegraphics[width=0.49\textwidth]{\detokenize{./figs/dish_example2.pdf}}
\begin{minted}[fontsize=\footnotesize,escapeinside=()]{bash}
mkfifo $t{0,1...}
curl $base/$y > $t0 & cat $t0 | split $t1 $t2 &
cat $t1 | xargs -n1 curl -s >$t3 & 
cat $t2 | xargs -n1 curl -s >$t4 &
...
cat $t9 | sort -rn > $t11 & cat $t10 | sort -rn > $t12 &
cat $t11 | eager > $t13 & cat $t12 | eager > $t14 &
sort -mrn $t13 $t14 > $t15 &
cat $t15 | head -n1 > $out1 &
wait $! && get-pids | xargs -n 1 kill -SIGPIPE
\end{minted}
\caption{
  \textbf{Output of \ttiny{pash --width=2} for Fig.~\ref{fig:example} (fragment).}
  \sys orchestrates parallel execution through named pipes, parallel operators, and custom runtime primitives---\eg \ttiny{eager}, \ttiny{split}, and \ttiny{get-pids}.
}
\vspace{-15pt}
\label{fig:example2}
\end{figure}

@@

\heading{Stateless Commands}
The first class, \sta, contains commands that operate on individual line elements of their input, without maintaining state across invocations.
These are commands that can be expressed as a purely functional \emph{map} or \emph{filter}---\eg \ttt{grep} filters out individual lines and \ttt{basename} removes a path prefix from a string.
They may produce multiple elements---\eg \ttt{tr} may insert {\sc NL} tokens---but always return empty output for empty input.
Workloads that use only stateless commands are trivial to parallelize:
  they do not require any synchronization to maintain correctness, nor caution about where to split inputs.
 % map :: (a -> b) -> [a] -> [b]

@@

\heading{Non-parallelizable Pure Commands}
The third class, \npu, contains commands that, while purely functional, cannot be parallelized.
This is because their internal state depends on prior state in the same pass in non-trivial ways. % should we say something about state machines?
For example, hashing commands such as \ttt{sha1sum} maintain complex state that has to be updated sequentially.
If parallelized on a single input, each stage would need to wait on the results of all previous stages, foregoing any parallelism benefits.

@@

Similarly, some pure commands (such as \ttt{sort} and \ttt{wc}) can be
parallelized using divide-and-conquer techniques. More formally, these
pure commands $f$ can be implemented as a combination of a function map
$m$ and an associative aggregate $agg$, satisfying the following equation:
%
\[
\forall x, x', s, f(x \cdot x', s) = agg(m(x, s), m(x', s), s)
\]
%
\noindent
This means that we get the same output by applying $f$ to a
  concatenation of two inputs $x, x'$ as by applying the aggregation function
  $agg$ to the outputs produced by applying map $m$ to each of $x$ and $x'$.
%% \km{We had also discussed here the case of
%%   operators that can be described as associative and commutative
%%   functions on the data (e.g., sum). In this case the parallelization
%%   has the flavor of data parallelism: one can split the input stream
%%   in an arbitrary way (e.g., randomly or using some other
%%   load-balancing scheme) and merge the outputs. For the more broad
%%   class of associative operations (e.g., sort), the formulation of
%%   this paragraph makes perfect sense. I just want to mention that this
%%   feels a lot more like \textbf{batching} than map-reduce-like data
%%   parallelization.} \kk{Thanks, I think you are right. Batching is a
%%   better name for it. In the end we didn't handle
%%   commutative-associative function (that are more like map-reduce) and
%%   so we changes the name to map-fold to not imply commutativity.}

@@

$T$ can be extended straightforwardly to nodes $v \in$ \pur,
implemented by a map-aggregate pair $(m, agg)$ as $ v(x_1 \cdot x_2
\cdots x_n, s) \Rightarrow agg(m(x_1, s), m(x_2, s), \ldots m(x_n,
s), s)$.
As long as the pair $(m, agg)$ meets the three invariants outlined earlier~\sx{ext}, 
$T$ can be shown to be behavior-preserving.
% \todo{It is shown in Fig.}\ref{fig:parallelization-transformation}.

\heading{Auxiliary Transformations}
%
\sys also performs a set of auxiliary transformations $t_{1-3}$ that are
depicted in \Cref{fig:auxiliary-transformations}. If a node has many
inputs, $t_1$ concatenates these inputs by inserting a \ttt{cat} node
to enable the parallelization transformations. In cases where a
parallelizable node has one input and is not preceded by a
concatenation, $t_2$ inserts a \ttt{cat} node that is preceded by its
inverse \ttt{split}, so that the concatenation can be commuted with
the node. Transformation $t_3$ inserts a relay node that
performs the identity transformation. Relay nodes can be useful for
monitoring and debugging, as well as for performance
improvements~\sx{optimizer}.
The additional relay nodes inserted by the auxiliary transformations affect the resulting % (task) 
  parallelism.

@@

% \subsection*{Acknowledgements}
% % Dumping people so that we don't forget
% We want to thank 
%   Andr\'e DeHon,
%   Ben Karel,
%   Caleb Stanford,
%   Thurston Dang,
%   Jean-S\'ebastien L\'egar\'e, and 
%   Nick Roessler.
% This material is based upon work supported by DARPA contract no. HR00112020013 (ALRSCC) and NSF award CCF 1763514.
